---
title: "Portfolio 3"
output: 
  html_document: 
    highlight: pygments
    theme: cerulean
---

```{r load-packages, message=FALSE}
library(tidyverse) 
library(carData)
data("WVS")
```

### The Goal

Building off of the prior portfolio piece, this portfolio piece seeks to dive deeper into public opinion. In the portfolio piece before, I got a good idea of the aparent trends in demographics of the data participants involving country, gender, college degree, and religion. Finally, I also investigated the general consensus regarding the main poverty question of whether respective government's were doing enough for the poverty issue. 

The goal of this portfolio piece is to understand the trends in public opinion of government poverty action by country, specifically in the United States. I am going to create a data visualization that showcases the survey results of the United States. Additionally, I will do this for all three other countries to understand how these countries differ in public opinion.

### The Data

We start with the data! The WVS data set is derrived from the carData package, viewed below.

```{r}
view(WVS)
nrow(WVS)
ncol(WVS)
```

The WVS data set comes from a survey collected in the United States, Australia, Norway, and Sweden. They received 5381 responses with most coming from Australian and American participants. 

The main variable of interest that the researchers intended to measure was public opinion of the respective government's action towards dealing with poverty in that country. Participants were asked the given question: “Do you think that what the government is doing for people in poverty in this country is about the right amount, too much, or too little?” to which they could respond: "Too Little", "About Right", "Too Much". Other variables accounted for are described below.

poverty
“Do you think that what the government is doing for people in poverty in this country is about the right amount, too much, or too little?” (ordered): Too Little, About Right, Too Much

religion
Member of a religion: no or yes.

degree
Held a university degree: no or yes.

country
Australia, Norway, Sweden, or USA.

age
in years.

gender
male or female.

This is a co-ed sample with almost an even distribution of male and female participants. Additionally, the data surveys variables such as having a college degree and being religious. Age is also accounted for in this data set and will be used in future portfolio pieces. 

### The Product

To start, here is the visualization produced in portfolio piece one that showcases the general trend of responses to the poverty question across all participants in all four countries.

```{r plot-poverty}
WVS %>%
  ggplot(data = WVS, mapping = aes(x = poverty)) +
  geom_bar()
```

As shown, the general majority believe their government is doing too little for the 

We are going to need to adjust our data set from the WVS data set. Firstly, I want to filter the data to only have responses from individuals based in the United States in order to guage public opinion in the US. I am going to use the filter() function.

```{r}
```


### The Interpretation